Statically Typed String Sanitation Inside a Python

### 2014-10-07 sanitation-psp14/sanitation-psp14-short.tex
Updated related work slightly
Final edits.

\subsection{The Language of Regular Strings}\label{sec:rs}
In this section, we define a minimal typed lambda calculus with regular string types called $\lambda_{RS}$. The syntax of $\lambda_{RS}$ is specified in Figure \ref{fig:glambdas}, its static semantics in Figure \ref{fig:slambdas} and its evaluation semantics in Figure \ref{fig:dlambdas}.\footnote{For convenience, a single sheet containing Figures \ref{fig:glambdas}-\ref{fig:tr} is available in the accompanying technical report \todo{TR}.}

@@

In our calculus, both forms of coercion are explicit. For safe coercions, it is often useful for the coercion to be performed implicitly. This can be seen as a form of  subtyping for regular string types \todo{cite?}. 
In practice, subtyping is crucial to the usability of our system due to the nature of the replacement operator.
For simplicity, we do not include this relation; however, subtyping for regular strings is present in previous treatments of our work \cite{fulton12,fulton13}. The implementation discussed in Section \ref{atlang} also provides subtyping between regular string types.

@@

\begin{prop}[Substitution Correspondence] \label{thm:substcorrespondence}
  If $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$ then $\lsubst{r}{s_1}{s_2} \in \lang{\lreplace{r}{r_1}{r_2}}$.
\end{prop}
\begin{proof}
The proposition follows from the definitions of $\textsf{subst}$ and $\textsf{lreplace}$; note that language substitutions over-approximate string substitutions.\todo{elaborate on this?}
\end{proof}

@@

We also need to establish a well-formedness lemma, which simply expresses the totality of the operations over regular expressions used as premises in the rules in Figure \ref{fig:slambdas}. 
\begin{lem}
  If $\sctx \vdash e : \stringin{r}$ then $r$ is a well-formed regular expression.
\end{lem}
\begin{proof}
  The only non-trivial case is S-T-Replace, which follows from Proposition \ref{thm:total}. \todo{what about lhead/ltail?}
\end{proof}

@@

% HosoyaPierce2002? 
Incorporating regular expressions into the type system is not novel. The XDuce system \cite{HosoyaVouillonPierce2000ICFP} checks XML documents against schema using regular expressions. 
Similarly, XHaskell \cite{xhaskell} focuses on XML documents.
We differ from this and related work in at least three ways:
\begin{itemize}
  \item Although our static replacement operation is definable in some languages with regular expression types, we are the first to expose this operation and connect the semantics of regular language replacement with the semantics of string substitution via a type safety and compilation correctness argument.
  \item We demonstrate that regular expression types are applicable to the web security domain, whereas previous work on regular expression types focused on XML.
  \item The underlying representation of regular strings is guaranteed to be a string, rather than a heavier implementation based on an encoding using functional datatypes.
  \item We work within an extensible type system.
\end{itemize}

@@

\section{Acknowledgements}
\todo{funding sources}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% END SECTION TWO -- THIS IS THE CONFLICT LINE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\subsection{The Language of Regular Strings}\label{sec:rs}
In this section, we define a minimal typed lambda calculus with regular string types called $\lambda_{RS}$. The syntax of $\lambda_{RS}$ is specified in Figure \ref{fig:glambdas}, its static semantics in Figure \ref{fig:slambdas} and its evaluation semantics in Figure \ref{fig:dlambdas}.\footnote{For convenience, a single sheet containing Figures \ref{fig:glambdas}-\ref{fig:tr} is available in the accompanying technical report \cite{psptr}.}

@@

In our calculus, both forms of coercion are explicit. For safe coercions, it is often useful for the coercion to be performed implicitly. This can be seen as a form of  subtyping for regular string types \cite{fulton13}. 
In practice, subtyping is crucial to the usability of our system due to the nature of the replacement operator.
For simplicity, we do not include this relation; however, subtyping for regular strings is present in previous treatments of our work \cite{fulton12,fulton13}. The implementation discussed in Section \ref{atlang} also provides subtyping between regular string types.

@@

\begin{prop}[Substitution Correspondence] \label{thm:substcorrespondence}
  If $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$ then $\lsubst{r}{s_1}{s_2} \in \lang{\lreplace{r}{r_1}{r_2}}$.
\end{prop}
\begin{proof}
The proposition follows from the definitions of $\textsf{subst}$ and $\textsf{lreplace}$; note that language substitutions over-approximate string substitutions.
\end{proof}

@@

We also need to establish a well-formedness lemma, which simply expresses the totality of the operations over regular expressions used as premises in the rules in Figure \ref{fig:slambdas}. 
\begin{lem}
  If $\sctx \vdash e : \stringin{r}$ then $r$ is a well-formed regular expression.
\end{lem}
\begin{proof}
  The only non-trivial cases are S-T-Replace and S-T-Concat. The first follows from Proposition \ref{thm:total}, and the other
  follows from the fact that lhead and ltail always produce well-formed expressions.
\end{proof}

@@

% HosoyaPierce2002? 
Incorporating regular expressions into the type system is not novel. The XDuce system \cite{HosoyaVouillonPierce2000ICFP} checks XML documents against schema using regular expressions. 
Similarly, XHaskell \cite{xhaskell} focuses on XML documents.
We differ from this and related work in at least three ways:
\begin{itemize}
  \item Although our static replacement operation is definable in some languages with regular expression types, we are the first to expose this operation and connect the semantics of regular language replacement with the semantics of string substitution via a type safety and compilation correctness argument.
  \item The underlying representation of regular strings is guaranteed to be a string, rather than a heavier implementation based on an encoding using functional datatypes.
  \item We demonstrate that regular expression types are applicable to the web security domain, whereas previous work on regular expression types focused on XML.
  \item We work within an extensible type system.
\end{itemize}

@@

\section{Acknowledgements}
This work was supported by supported by the National Security Agency
lablet contract \#H98230-14-C-0140
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% END SECTION TWO -- THIS IS THE CONFLICT LINE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
